Reproduce analyses reported in: *Phenotypic evolution is more constrained in simpler food webs*

### 2019-01-16 analyses/reproduce_analyses.Rmd
many minor formatting edits to make it easier to understand the code and results.

---
title: "Reproduce analyses reported in: *Phenotypic evolution is more constrained in simpler food webs*"
author: "Matthew A. Barbour"
date: "`r Sys.Date()`"
output: github_document
fontsize: 12pt
---

```{r setup, include=FALSE}

## Load required libraries ----
library(car)        # for homogeneity of variance test
library(tidyverse)  # for managing data
library(cowplot)    # pretty default ggplots
library(broom)      # for tidying multiple linear models
library(visreg)     # for quick visualizations of model effects
library(viridis)    # for color palette
library(lme4)       # for generalized linear mixed models
library(piecewiseSEM) # for calculating model R-square
library(latex2exp)  # using Latex notation for axes labels
library(stringr)
library(MVN)        # assessing multivariate normality assumption

## Load and manage data ----

gall_selection.df <- read_csv("../data/gall_selection_data.csv") %>%
  # convert appropriate variables to characters instead of integers
  mutate(Plant_Position = as.character(Plant_Position),
         Gall_Number = as.character(Gall_Number)) %>%
  unite(Gall_ID, Gall_Number, Gall_Letter, remove = FALSE) %>%

  # subset data for analysis
  filter(phenology == "early", Location == "tree",
         platy > 0 | ectos > 0 | pupa > 0) %>%                  # eliminate unknown sources of mortality
  filter(Treatment.focus == "Ectoparasitoid exclusion" & ectos < 1 | Treatment.focus == "Control") %>% # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
  mutate(Foodweb = ifelse(Treatment.focus=="Control","Complex","Simple"),
         gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg_parasitoid = as.numeric(ifelse(platy > 0, 1, 0)),
         # transform and scale traits to approximate multivariate normal assumption
         sc.Diam = as.numeric(scale(Gall_Height_mm)), 
         sc.log.Clutch = as.numeric(scale(log(gall_individuals))), 
         sc.sqrt.Pref = as.numeric(scale(sqrt(Density_per_100_shoots))))

## Custom functions to avoid repetitive code ----

## Transform logit to probability
inverse_logit <- function(x) exp(x)/(1+exp(x))

## Get data for plotting bootstrapped estimates in Figures 2, 3, and 4 of main text
bootstrap_fitness <- function(logistic_model, newdata, bootstraps, intervals=c(0.025,0.975)){

@@

## Number of bootstrap replicates
n_boots_analysis <- 1000
n_boots_plots <- 100

## Get confidence intervals from bootstrapped samples
conf.high <- function(x) quantile(x, probs = 0.975)
conf.low <- function(x) quantile(x, probs = 0.025)

## Color-blind friendly palette: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/

# order -> orange, light blue, green, yellow, dark blue, red, pink
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000", "#999999")
complex_color <- cbPalette[6] # 1
simple_color <- cbPalette[5] # 2
treatment_colors <- c(complex_color, simple_color)

@@

# Evaluating assumption of multivariate normality

We used graphical checks to evaluate whether our transformations of trait values resulted in a multivariate normal distribution. Figure S\ref{fig:Univariate_histograms} shows that our transformations resulted in approximately normal distributions for each phenotypic trait. Note also that in the multivariate quantile-quantile (Q-Q) plot, most points fall along the expected line (fig. S\ref{fig:Multivariate_QQ}), suggesting that our transformations provide a reasonable approximation of a multivariate normal distribution.

@@

```{r Multivariate_QQ, fig.cap="\\label{fig:Multivariate_QQ}Multivariate quantile-quantile (Q-Q) plot to assess deviations from multivariate normality (black line)."}
mvn_multivariate_qq <- mvn(select(gall_selection.df, sc.Diam, sc.log.Clutch, sc.sqrt.Pref), multivariatePlot = "qq")
```

# Effect of food-web treatment on trait-fitness relationships and selection gradients

@@

# Accuracy of confidence intervals is not a priority here, so we just use the asymptotic estimates rather than parametric bootstrapping.
biased_foodweb_confint <- tidy(biased_foodweb_model, conf.int=TRUE) %>% filter(group=="fixed")

# Print table
knitr::kable(biased_foodweb_confint)
```

We gather regression coefficients from the full and reduced model to display trait-fitness relationships. For chamber diameter, we subtracted the coefficient due to bias to better approximate the trait-fitness relationship. We also used the bootstrapped samples to calculate the differences in estimates between treatments (labelled **Diff**). The code for tidying this data is not displayed, but is available in the 'reproduce_analyses.Rmd'.

@@

# Tidy up estimates
tidy_foodweb_alphas <- foodweb_alphas %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"),
         coefficient_type = ifelse(term=="(Intercept)","Mean fitness",
                                   ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
                                          ifelse(grepl(":", .$term)==TRUE, "Correlational","Linear"))))
```

Table of trait-fitness relationships:

```{r Table of food web coefficients}
knitr::kable(tidy_foodweb_alphas)
```

Plot these trait-fitness relationships:

```{r Plot food-web coefficients}

@@

# Plot mean fitness
plot_foodweb_mean_fitness <- tidy_foodweb_alphas %>%
  filter(type!="Diff", term=="(Intercept)") %>%
  ggplot(., aes(x=type)) +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  geom_text(data=filter(tidy_foodweb_alphas, type=="Diff", term=="(Intercept)"), aes(y=0.8, x=1.5, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="larval survival", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  xlab("Food web") +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels) +
  scale_color_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels)

@@

@@

Table of selection gradients.

```{r Table of food web selection gradients}
knitr::kable(tidy_foodweb_grads)
```

Plot of selection gradients.

```{r Plot food-web gradients}

@@

# Partitioning the contribution of egg and larval parasitoids to selection gradients

@@

```{r Egg vs. larval ptoid model, cache=TRUE}
# excludes cases of egg-parasitism from Complex food web
egglarval_df <- filter(gall_selection.df, Foodweb == "Simple" | Foodweb == "Complex" & platy < 1)

@@

Parametric bootstrapping to estimate confidence intervals

@@

Table of trait-fitness relationships for egg and larval parasitoids.

```{r Tabel of egg vs. larval coefficients}
knitr::kable(tidy_egglarval_alphas)
```

Plot of trait-fitness relationships for egg and larval parasitoids.

```{r Plot egglarval coefficients}

@@

# Plot mean fitness
plot_egglarval_mean_fitness <- tidy_egglarval_alphas %>%
  filter(type!="Diff", term=="(Intercept)") %>%
  ggplot(., aes(x=type)) +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  geom_text(data=filter(tidy_egglarval_alphas, type=="Diff", term=="(Intercept)"), aes(y=0.9, x=1.5, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="larval survival", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  scale_x_discrete(name="Parasitoid", breaks=c("Complex","Simple"), labels=c("Larval","Egg")) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels) +
  scale_color_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels)

@@

Calculate selection gradients from trait-fitness relationships.

@@

Table of selection gradients imposed by egg and larval parasitoids.

```{r Egg vs. larval selection gradients}
knitr::kable(tidy_egglarval_grads)
```

Plot of selection gradients imposed by egg and larval parasitoids.

```{r Plot egg vs. larval ptoid gradients}

@@

# Plot mean fitness
plot_foodwebegglarval_mean_fitness <- tidy_foodwebegglarval_alphas %>%
  filter(type!="Diff", term=="(Intercept)") %>%
  ggplot(., aes(x=type)) +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  scale_y_continuous(name="larval survival", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  scale_x_discrete(name="Parasitoids", breaks=c("Complex","Simple","Larval"), labels=c("Egg + Larval","Egg","Larval")) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(6,5,4)], labels=legend_foodwebegglarval_labels) +
  scale_color_manual(values=cbPalette[c(6,5,4)], labels=legend_foodwebegglarval_labels)

@@

# Reproduce Figure 2 in main text of manuscript.

@@

```{r Univariate-Fitness-Landscapes, fig.cap="Selection gradients acting on gall traits in complex vs. simple food webs. Each panel corresponds to a different gall trait: gall diameter (A); clutch size (B); and Oviposition preference (C). Solid lines represent the estimated gradients in complex (orange) and simple (blue) food webs. Transparent lines represent bootstrapped replicates (n=100) to show the uncertainty in estimated gradients. Note that only 100 bootstraps are displayed here, but that inferences are based on 1,000 bootstrapped samples."}

@@

# Replot univariate adaptive landscapes in terms of relative fitness

```{r Plot Univariate RF Landscapes}

@@

## Female pref

@@

# Reproduce Figure 3 presented in main text.

@@

```{r Figure-Multivariate-Landscapes}

@@

# Calculating Slope and Curvature of the Adaptive Landscape

@@

## Create beta-matrix for complex food-web treatment
complex_betas_df <- tidy_foodweb_grads %>%
  filter(type=="Complex", gradient_type=="Directional") %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)
complex_betas_df$term <- factor(complex_betas_df$term, levels=c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref"), labels=c("Diam","Clutch","Pref"), order=TRUE)
rownames(complex_betas_df) <- complex_betas_df$term

@@

## Create gamma-matrix for complex food-web treatment
complex_gammas_df <- tidy_foodweb_grads %>%
  filter(type=="Complex", gradient_type %in% c("Quadratic","Correlational")) %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)
complex_gammas_df$term <- factor(complex_gammas_df$term, levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"), labels=c("Diam^2","Clutch^2","Pref^2","Diam:Clutch","Diam:Pref","Clutch:Pref"), order=TRUE)
rownames(complex_gammas_df) <- complex_gammas_df$term

@@

## Create matrix of Betas in simple food-web treatment
simple_betas_df <- tidy_foodweb_grads %>%
  filter(type=="Simple", gradient_type=="Directional") %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)
simple_betas_df$term <- factor(simple_betas_df$term, levels=c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref"), labels=c("Diam","Clutch","Pref"), order=TRUE)
rownames(simple_betas_df) <- simple_betas_df$term

@@

## Create gamma-matrix for simple food-web treatment
simple_gammas_df <- tidy_foodweb_grads %>%
  filter(type=="Simple", gradient_type %in% c("Quadratic","Correlational")) %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)
simple_gammas_df$term <- factor(simple_gammas_df$term, levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"), labels=c("Diam^2","Clutch^2","Pref^2","Diam:Clutch","Diam:Pref","Clutch:Pref"), order=TRUE)
rownames(simple_gammas_df) <- simple_gammas_df$term

@@

## Calculate Curvature of the Fitness landscape
complex_curvature <- complex_gammas - complex_betas %*% t(complex_betas)
simple_curvature <- simple_gammas - simple_betas %*% t(simple_betas)
```

@@

# Estimating selection on the egg parasitoid *Platygaster*

@@

# convert "gall_survival" to egg parasitoid survival. Note that both Iteomyia pupa and larva parasitoids result in 0. We do not change the name of the response variable to make clear that we are using the same statistical models as for Iteomyia.
eggegg_df <- mutate(gall_selection.df, gall_survival = ifelse(egg_parasitoid==1,1,0))

eggegg_model <- update(foodweb_model, data=eggegg_df)

biased_eggegg_df <- eggegg_df %>%
  group_by(Foodweb, Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  ungroup()

biased_eggegg_model <- update(biased_foodweb_model, data=biased_eggegg_df)
```

@@

@@

```{r Bootstrap linear egg vs. egg model, cache=TRUE, include=FALSE}
boot_linear_eggegg_model <- bootMer(linear_eggegg_model, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32, seed=34)
```

```{r Egg vs. egg ptoid alpha coefficients, include=FALSE}
eggegg_bind <- bind_cols(
  as_tibble(boot_linear_eggegg_model$t),
  select(as_tibble(boot_eggegg_model$t), `FoodwebComplex:I(sc.Diam^2)`:`FoodwebSimple:sc.log.Clutch:sc.sqrt.Pref`) # only retain nonlinear coefficients from full model
)

@@

```{r Egg vs. egg ptoid coefficients}
knitr::kable(tidy_eggegg_alphas)
```

```{r Plot egg vs. egg ptoid coefficients}

>>>>

---
title: "Reproduce analyses reported in: *Phenotypic evolution is more constrained in simpler food webs*"
author: "Matthew A. Barbour"
date: "`r Sys.Date()`"
output:  
  prettydoc::html_pretty:
    theme: architect
    highlight: github
fontsize: 11pt
---

```{r setup}
knitr::opts_chunk$set(messages=FALSE, warning=FALSE, error=FALSE)

# Load required libraries ----
library(car)        # for homogeneity of variance test
library(tidyverse)  # for managing data
library(cowplot)    # pretty default ggplots
library(broom)      # for tidying multiple linear models
#library(visreg)     # for quick visualizations of model effects
library(viridis)    # for color palette
library(lme4)       # for generalized linear mixed models
#library(piecewiseSEM) # for calculating model R-square
library(latex2exp)  # using Latex notation for axes labels
library(stringr)
library(MVN)        # assessing multivariate normality assumption

# Load and manage data ----

gall_selection.df <- read_csv("../data/gall_selection_manuscript_dataset.csv") %>%
  # convert appropriate variables to characters instead of integers
  mutate(Plant_Position = as.character(Plant_Position),
         Gall_Number = as.character(Gall_Number)) %>%
  unite(Gall_ID, Gall_Number, Gall_Letter, remove = FALSE) %>%

  # subset data for analysis
  filter(#phenology == "early", Location == "tree",
         egg.ptoid > 0 | larval.ptoid > 0 | pupa > 0) %>%                  # eliminate unknown sources of mortality
  filter(Treatment == "Ectoparasitoid exclusion" & larval.ptoid < 1 | Treatment == "Control") %>% # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
  mutate(Foodweb = ifelse(Treatment=="Control","Complex","Simple"),
         gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg.ptoid = as.numeric(ifelse(egg.ptoid > 0, 1, 0)), # there were a couple occassions where two egg parasitoids were found in a single chamber. I set these to 1 to permit an analysis with a binomial model.
         # transform and scale traits to approximate multivariate normal assumption
         sc.Diam = as.numeric(scale(chamber.diameter_mm)), 
         sc.log.Clutch = as.numeric(scale(log(clutch.size_individuals))), 
         sc.sqrt.Pref = as.numeric(scale(sqrt(oviposition.preference_individuals.per.100.shoots))))

# Custom functions to avoid repetitive code ----

# Transform logit to probability
inverse_logit <- function(x) exp(x)/(1+exp(x))

# Get data for plotting bootstrapped estimates in Figures 2, 3, and 4 of main text
bootstrap_fitness <- function(logistic_model, newdata, bootstraps, intervals=c(0.025,0.975)){

@@

# Number of bootstrap replicates
n_boots_analysis <- 100
n_boots_plots <- 10

# Get confidence intervals from bootstrapped samples
conf.high <- function(x) quantile(x, probs = 0.975)
conf.low <- function(x) quantile(x, probs = 0.025)

# Color-blind friendly palette: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/; order -> orange, light blue, green, yellow, dark blue, red, pink
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000", "#999999")
complex_color <- cbPalette[6] 
simple_color <- cbPalette[5] 
treatment_colors <- c(complex_color, simple_color)

@@

\

## Evaluating assumption of multivariate normality

We used graphical checks to evaluate whether our transformations of trait values resulted in a multivariate normal distribution. The histograms below show that our transformations resulted in approximately normal distributions for each phenotypic trait. Note also that in the multivariate quantile-quantile (Q-Q) plot, most points fall along the expected line, suggesting that our transformations provide a reasonable approximation of a multivariate normal distribution.

@@

```{r Multivariate_QQ, fig.cap="Multivariate quantile-quantile (Q-Q) plot to assess deviations from multivariate normality (black line)."}
mvn_multivariate_qq <- mvn(select(gall_selection.df, sc.Diam, sc.log.Clutch, sc.sqrt.Pref), multivariatePlot = "qq")
```

## Gall trait-fitness relationships: effect of food-web treatment

@@

# Accuracy of confidence intervals is not a priority here, so we just use the asymptotic estimates rather than parametric bootstrapping.
biased_foodweb_confint <- tidy(biased_foodweb_model, conf.int=TRUE) %>% filter(group=="fixed")
```

```{r biased foodweb table, echo=FALSE}
# Print table
biased_foodweb_confint %>%
  select(term, estimate, conf.low, conf.high) %>%
  filter(term %in% c("FoodwebComplex:sc.Diam","FoodwebSimple:sc.Diam")) %>%
  knitr::kable(., caption = "Estimates of bias in trait-fitness relationship of chamber diameter in each food-web treatment.")
```

We gather regression coefficients from the full and reduced model to display trait-fitness relationships. For chamber diameter, we subtracted the coefficient due to bias to better approximate the trait-fitness relationship. We also used the bootstrapped samples to calculate the differences in estimates between treatments (labelled **Diff**).

@@

# Tidy up estimates
tidy_foodweb_alphas <- foodweb_alphas %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"),
         coefficient_type = ifelse(term=="(Intercept)","Mean fitness",
                                   ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
                                          ifelse(grepl(":", .$term)==TRUE, "Correlational","Linear"))))
```
\

```{r Table of food web coefficients, echo=FALSE}
tidy_foodweb_alphas %>%
  filter(type != "Diff") %>%
  select(term, coefficient_type, type, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%
 knitr::kable(., caption = "Table of mean fitness and trait-fitness relationships in each food-web treatment.")
```

\

```{r Table of differences in food web coefficients, echo=FALSE}
tidy_foodweb_alphas %>%
  filter(type == "Diff") %>%
  select(term, coefficient_type, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%
 knitr::kable(., caption = "Table of differences in mean fitness and trait-fitness relationships between food-web treatments.")
```

\

The table is good for details, but it is easier to see the results as a figure.

```{r Plot food-web coefficients, fig.cap="Plot of mean fitness differences between treatments and trait-fitness relationships. Asterisks denote significant differences (*P* < 0.05) between treatments."}

@@

# Plot mean fitness
plot_foodweb_mean_fitness <- tidy_foodweb_alphas %>%
  filter(type!="Diff", term=="(Intercept)") %>%
  ggplot(., aes(x=type)) +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  geom_text(data=filter(tidy_foodweb_alphas, type=="Diff", term=="(Intercept)"), aes(y=0.8, x=1.5, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Larval survival", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  xlab("Food web") +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels) +
  scale_color_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels)

@@

\

## Gall trait selection gradients

@@

\

Reproduce numbers reported in **Table 1** of the manuscript.

```{r Table of food web selection gradients, echo=FALSE}
tidy_foodweb_grads %>%
  filter(type != "Diff") %>%
  select(term, gradient_type, type, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%
knitr::kable(., caption = "Table of selection gradients acting on gall traits in each food-web treatment.")
```

\

```{r Table of differences in selection gradients between food-web treatments, echo=FALSE}
tidy_foodweb_grads %>%
  filter(type == "Diff") %>%
  select(term, gradient_type, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%
  knitr::kable(., caption = "Table of differences in selection gradients acting on gall traits between food-web treatment.")
```

\

Now visualize the table output as a figure.

```{r Plot food-web gradients, fig.cap="Plot of selection gradients in each food-web treatment. Asterisks denote significant differences (*P* < 0.05) between food-web treatments."}

@@

\

## Gall trait-fitness relationships: contribution of egg and larval parasitoids

@@

```{r Egg vs. larval ptoid model, cache=TRUE}
# excludes cases of egg-parasitism from Complex food web
egglarval_df <- filter(gall_selection.df, Foodweb == "Simple" | Foodweb == "Complex" & egg.ptoid < 1)

@@

@@

\

```{r Table of egg vs. larval coefficients, echo=FALSE}
tidy_egglarval_alphas %>%
  filter(type != "Diff") %>%
  select(term, coefficient_type, type, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%   knitr::kable(., caption = "Table of gall trait-fitness relationships imposed by egg and larval parasitoids. Note that type = Simple = Egg parasitoid and Complex = Larval parasitoid.")
```

```{r Table of differences in egg vs. larval coefficients, echo=FALSE}
tidy_egglarval_alphas %>%
  filter(type == "Diff") %>%
  select(term, coefficient_type, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%   knitr::kable(., caption = "Table of differences in gall trait-fitness relationships between egg and larval parasitoids.")
```

See the table results in figure form.

```{r Plot egglarval coefficients, fig.cap="Trait-fitness relationships for egg and larval parasitoids. Asterisks denote significant differences (*P* < 0.05) between parasitoid guilds."}

@@

# Plot mean fitness
plot_egglarval_mean_fitness <- tidy_egglarval_alphas %>%
  filter(type!="Diff", term=="(Intercept)") %>%
  ggplot(., aes(x=type)) +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  geom_text(data=filter(tidy_egglarval_alphas, type=="Diff", term=="(Intercept)"), aes(y=0.9, x=1.5, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Larval survival", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  scale_x_discrete(name="Parasitoid", breaks=c("Complex","Simple"), labels=c("Larval","Egg")) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels) +
  scale_color_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels)

@@

## Selection gradients acting on gall traits due to egg and larval parasitoids

@@

\

```{r Egg vs. larval selection gradients, echo=FALSE}
tidy_egglarval_grads %>%
  filter(type != "Diff") %>%
  select(term, gradient_type, type, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%
knitr::kable(., caption = "Table of selection gradients imposed by egg and larval parasitoids.")
```

\

```{r Table of differences in selection gradients between egg and larval parasitoids, echo=FALSE}
tidy_egglarval_grads %>%
  filter(type == "Diff") %>%
  select(term, gradient_type, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%
  knitr::kable(., caption = "Table of differences in selection gradients imposed by egg and larval parasitoids.")
```

\

See table results in figure form.

```{r Plot egg vs. larval ptoid gradients, fig.cap="Selection gradients imposed by egg and larval parasitoids. Asterisks denote significant differences (*P* < 0.05) between parasitoid guilds."}

@@

# Plot mean fitness
plot_foodwebegglarval_mean_fitness <- tidy_foodwebegglarval_alphas %>%
  filter(type!="Diff", term=="(Intercept)") %>%
  ggplot(., aes(x=type)) +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  scale_y_continuous(name="Larval survival", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  scale_x_discrete(name="Parasitoids", breaks=c("Complex","Simple","Larval"), labels=c("Egg + Larval","Egg","Larval")) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(6,5,4)], labels=legend_foodwebegglarval_labels) +
  scale_color_manual(values=cbPalette[c(6,5,4)], labels=legend_foodwebegglarval_labels)

@@

\

## Reproduce Figure 2 in manuscript.

@@

```{r Univariate-Fitness-Landscapes, fig.cap="Adaptive landscapes of gall phenotypes in complex and simple food webs."}

@@

\

## Replot adaptive landscapes of gall traits in terms of relative fitness

```{r Plot Univariate RF Landscapes, fig.cap="Adaptive landscapes of gall phenotypes in terms of relative fitness."}

@@

# Female pref

@@

## Reproduce Figure 3 in manuscript

@@

```{r Figure-Multivariate-Landscapes, fig.cap="Multivariate adaptive landscapes of gall phenotypes in complex and simple food webs. Dotted lines denote fitness for the mean phenotype. Arrows represent the vector of directional selection gradients acting on each phenotype."}

@@

## Slope and curvature of the adaptive landscape of gall traits

@@

# Create beta-matrix for complex food-web treatment
complex_betas_df <- tidy_foodweb_grads %>%
  filter(type=="Complex", gradient_type=="Directional") %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)
complex_betas_df$term <- factor(complex_betas_df$term, levels=c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref"), labels=c("Diam","Clutch","Pref"), order=TRUE)
rownames(complex_betas_df) <- complex_betas_df$term

@@

# Create gamma-matrix for complex food-web treatment
complex_gammas_df <- tidy_foodweb_grads %>%
  filter(type=="Complex", gradient_type %in% c("Quadratic","Correlational")) %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)
complex_gammas_df$term <- factor(complex_gammas_df$term, levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"), labels=c("Diam^2","Clutch^2","Pref^2","Diam:Clutch","Diam:Pref","Clutch:Pref"), order=TRUE)
rownames(complex_gammas_df) <- complex_gammas_df$term

@@

# Create matrix of Betas in simple food-web treatment
simple_betas_df <- tidy_foodweb_grads %>%
  filter(type=="Simple", gradient_type=="Directional") %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)
simple_betas_df$term <- factor(simple_betas_df$term, levels=c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref"), labels=c("Diam","Clutch","Pref"), order=TRUE)
rownames(simple_betas_df) <- simple_betas_df$term

@@

# Create gamma-matrix for simple food-web treatment
simple_gammas_df <- tidy_foodweb_grads %>%
  filter(type=="Simple", gradient_type %in% c("Quadratic","Correlational")) %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)
simple_gammas_df$term <- factor(simple_gammas_df$term, levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)","sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"), labels=c("Diam^2","Clutch^2","Pref^2","Diam:Clutch","Diam:Pref","Clutch:Pref"), order=TRUE)
rownames(simple_gammas_df) <- simple_gammas_df$term

@@

# Calculate Curvature of the Fitness landscape
complex_curvature <- complex_gammas - complex_betas %*% t(complex_betas)
simple_curvature <- simple_gammas - simple_betas %*% t(simple_betas)
```

@@

\

## Trait-fitness relationships of the extended phenotype of the egg parasitoid *Platygaster* sp.

@@

# convert "gall_survival" to egg parasitoid survival. Note that both Iteomyia pupa and larva parasitoids result in 0. We do not change the name of the response variable to make clear that we are using the same statistical models as for Iteomyia.
eggegg_df <- mutate(gall_selection.df, gall_survival = ifelse(egg.ptoid==1,1,0))

# fit trait-fitness relationship using the same model structure
eggegg_model <- update(foodweb_model, data=eggegg_df)

# subset data for quantifying biased selection on chamber diameter
biased_eggegg_df <- eggegg_df %>%
  group_by(Foodweb, Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  ungroup()

# quantify bias
biased_eggegg_model <- update(biased_foodweb_model, data=biased_eggegg_df)
```

Use parametric bootstrapping to calculating 95% confidence intervals of trait-fitness relationships.

@@

Remove all higher-order terms to quantify linear trait-fitness relationships.

@@

```{r Bootstrap linear egg vs. egg model, cache=TRUE}
boot_linear_eggegg_model <- bootMer(linear_eggegg_model, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32, seed=34)
```

Gather regression coefficients to display trait-fitness relationships.

```{r Egg vs. egg ptoid alpha coefficients}
eggegg_bind <- bind_cols(
  as_tibble(boot_linear_eggegg_model$t),
  select(as_tibble(boot_eggegg_model$t), `FoodwebComplex:I(sc.Diam^2)`:`FoodwebSimple:sc.log.Clutch:sc.sqrt.Pref`) # only retain nonlinear coefficients from full model
)

@@

\

```{r Egg ptoid coefficients in each food-web treatment, echo=FALSE}
tidy_eggegg_alphas %>%
  filter(type != "Diff") %>%
  select(term, coefficient_type, type, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%
knitr::kable(., caption = "Table of trait-fitness relationships of extended phenotype of egg parasitoids in complex and simple food webs.")
```

\

```{r Table of differences in egg parasitoid coefficients between food-web treatments, echo=FALSE}
tidy_eggegg_alphas %>%
  filter(type == "Diff") %>%
  select(term, coefficient_type, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 3), conf.low = round(conf.low, 3), conf.high = round(conf.high, 3)) %>%
  knitr::kable(., caption = "Table of differences in trait-fitness relationships of egg parasitoids between food-web treatment.")
```

\

Visualize tables in figure form.

```{r Plot egg vs. egg ptoid coefficients, fig.cap="Trait-fitness relationships of egg parasitoid's extended phenotype in complex and simple food webs. Asterisks denote significant differences (*P* < 0.05) between food-web treatments."}
